\pdfbookmark[1]{Additional Information}{additional_information}

\pdfbookmark[2]{Ethics Declarations}{ethics_declarations}                    
\section*{Ethics Declarations}
This study used only synthetic data and publicly available datasets. No human subjects or animal experiments were conducted specifically for this research. All analyses were performed in accordance with relevant guidelines and regulations.
\label{ethics declarations}

\pdfbookmark[2]{Contributors}{author_contributions}                    
\section*{Author Contributions}
Y.W. conceptualized the study, designed and implemented the gPAC framework, performed all analyses, and wrote the manuscript.
\label{author contributions}

\pdfbookmark[2]{Acknowledgments}{acknowledgments}                    
\section*{Acknowledgments}
We thank the open-source community for valuable feedback during development. Computational resources were provided by [Institution]. We acknowledge the developers of PyTorch, NumPy, and TensorPAC for creating the foundational tools that enabled this work.
\label{acknowledgments}

\pdfbookmark[2]{Declaration of Interests}{declaration_of_interest}                    
\section*{Declaration of Interests}
The authors declare that they have no competing interests.
\label{declaration of interests}

\pdfbookmark[2]{Inclusion and Diversity Statement}{inclusion_and_diversity_statement}        
\section*{Inclusion and Diversity Statement}
We support inclusive, diverse, and equitable conduct of research. The gPAC framework is designed to be accessible to researchers regardless of their computational background, with comprehensive documentation and examples in multiple languages.
\label{inclusion and diversity statement}

\pdfbookmark[2]{Declaration of Generative AI in Scientific Writing}{declaration_of_generative_ai}
\section*{Declaration of Generative AI in Scientific Writing}
During the preparation of this work, the authors used Claude (Anthropic) to assist with code documentation and manuscript editing. After using this tool, the authors reviewed and edited the content as needed and take full responsibility for the content of the publication.
\label{declaration of generative ai in scientific writing}